Near Optimal Behavior via Approximate State Abstraction

### 2016-05-21 paper/introduction.tex
Changed classes of abstraction to types of abstraction

% Summary of the four abstractions.
We support this thesis by describing four different classes of approximate abstraction functions that preserve near-optimal behavior by aggregating states on different criteria: $\epQ$, on similar optimal $Q$-values, $\epM$, on similarity of rewards and transitions, $\epB$, on similarity of a Boltzmann distribution over optimal $Q$-values, and $\epMu$, on similarity of a multinomial distribution over optimal $Q$-values. Furthermore, we empirically demonstrate the relationship between the degree of compression and error incurred on a variety of \acp{MDP}.

>>>>

% Summary of the four abstractions.
We support this thesis by describing four different types of approximate abstraction functions that preserve near-optimal behavior by aggregating states on different criteria: $\epQ$, on similar optimal $Q$-values, $\epM$, on similarity of rewards and transitions, $\epB$, on similarity of a Boltzmann distribution over optimal $Q$-values, and $\epMu$, on similarity of a multinomial distribution over optimal $Q$-values. Furthermore, we empirically demonstrate the relationship between the degree of compression and error incurred on a variety of \acp{MDP}.

### 2016-05-21 paper/theory_results.tex
Added QED to end of lemma proofs
Changed wedge for AND
Changed classes of abstraction to types of abstraction

We now define classes of abstraction based on functions of state--action pairs.
\bdefn{$\ep_f$}
Given a function $f: \mathcal{S}_G \times \mathcal{A} \rightarrow \mathbb{R}$ and a fixed non-negative $\varepsilon \in \mathbb{R}$, we define $\ep_f$ as a type of approximate state-aggregation function that satisfies the following for any two ground states $s_1$ and $s_2$: 
\begin{equation}
\label{eq:phi_f}
\ep_f(s_1) = \ep_f(s_2) \rightarrow \forall_a \left|f(s_1, a) - f(s_2, a)\right| \leq \varepsilon.
\end{equation}
\edefn

@@

% Theorem 1
\begin{thm}
There exist at least four classes of approximate state-aggregation functions, $\epQ$, $\epM$, $\epB$ and $\epMu$, for which the optimal policy in the abstract \ac{MDP}, applied to the ground \ac{MDP}, has suboptimality bounded polynomially in $\frac{1}{\varepsilon}$:
\begin{equation}
\forall_{s \in \mathcal{S}_G} V_G^{\pi_G^*}(s) - V_G^{\pi_{GA}}(s) \leq poly \left (\frac{1}{\varepsilon} \right).
\end{equation}
\end{thm}

@@

Note that as $t \rightarrow \infty$, $\sum_{i=0}^{t}\gamma^i \frac{2\varepsilon}{1-\gamma} \rightarrow \frac{2\varepsilon}{(1-\gamma)^2}$ by the sum of infinite geometric series and $\pi_{A,t}(s) \rightarrow \pi_{GA}$.
Thus, we conclude Lemma~\ref{lma:Q*}.
%\begin{equation*}
%\forall_{s_G \in \mathcal{S}_g} V_G(s_G) \leq  V_G^{\pi_{GA}}(s_G) + \frac{2\varepsilon}{(1-\gamma)^2}
%\end{equation*}
%\begin{equation*}
%\forall_{s_G \in \mathcal{S}_g} \left | V_G(s_G) - V_G^{\pi_{GA}}(s_G) \right | \leq  \frac{2\varepsilon}{(1-\gamma)^2}
%\end{equation*}

@@

Now, consider an approximate version of \citet{li2006towards}'s $\phi_{model}$, where states are aggregated together when their rewards and transitions are within $\varepsilon$.
\bdefn{$\epM$}
We let $\epM$ define a type of abstraction that, for fixed $\varepsilon$, satisfies:
\begin{multline}
\epM(s_1) = \epM(s_2) \rightarrow \\
\forall_a \left| \mcR_G(s_1, a) - \mcR_G(s_2, a)\right| \leq \varepsilon\; \wedge\ \\
\forall_{s_A \in \mcS_A} \left|\sum_{{s_G}' \in G(s_A)} \left[\mcT_G(s_1, a, {s_G}') - \mcT_G(s_2, a,{s_G}')\right] \right| \leq \varepsilon.
\end{multline}
\edefn

@@

Let $B$ be the maximum $Q$-value difference between any pair of ground states in the same abstract state for $\epM$:
\begin{equation*}
B = \max_{s_A, s_1, s_2, a}  |Q_G(s_1, a) - Q_G(s_2, a)|,
\end{equation*}
where $s_A \in \mcS_G$ and $s_1, s_2 \in G(s_A)$. Since difference of rewards is bounded by $\varepsilon$:
% Removed for space
%\begin{multline*}
%B=\max_{s_A, s_1, s_2, a}      \biggl|\mcR_G(s_1, a) - \mcR_G(s_2, a)\ + \gamma \sum_{{s_G}' \in \mcS_G} \\
%\biggl[(\mcT_G(s_1,a,{s_G}') -\mcT_G(s_2, a, {s_G}'))\max_{a'}Q_G({s_G}', a')\biggr]\biggr|
%\end{multline*}
\begin{multline*}
B\leq \varepsilon + \gamma \sum_{s_A \in \mcS_A}\sum_{{s_G}' \in G(s_A)} \biggl[(T_G(s_1, a, {s_G}')\ -\\ 
T_G(s_2, a, {s_G}')) \max_{a'}Q_G({s_G}', a') \biggr].
\end{multline*}
By similarity of transitions under $\epM$:
\begin{align*}
B \leq \varepsilon + \gamma \textsc{QMax} \sum_{s_A \in \mcS_A} \varepsilon \leq \varepsilon + \gamma|\mcS_G|\varepsilon \textsc{QMax}.
\end{align*}
Since \textsc{QMax} $\leq \frac{\textsc{RMax}}{1-\gamma}$, and we defined $\textsc{RMax} = 1$:
\begin{equation*}
B \leq \frac{\varepsilon + \gamma(|\mcS_G| - 1) \varepsilon}{1-\gamma}.
\end{equation*}
Since the $Q$-values of ground states grouped under $\epM$ are strictly less than $B$, we can understand $\epM$ as a type of $\epQ$ with $\varepsilon = B$. Applying Lemma \ref{lma:Q*} yields Lemma \ref{lma:model}.
\subsection{Boltzmann over Optimal Q: $\epB$}
\label{sec:boltz}

@@

By the approximation in Equation~\ref{eq:e_to_x_approx} and the assumption in Equation~\ref{eq:bolt_denom}:
\begin{align*}
\left|\frac{1 + Q_G(s_1,a) + \delta_1}{\sum_j e^{Q_G(s_1,a_j)}} - \frac{1 + Q_G(s_2,a) + \delta_2}{\sum_j e^{Q_G(s_1,a_j)} \pm k\varepsilon}\right| \leq \varepsilon
\end{align*}
It follows that:
\begin{equation}
\left|Q_G(s_1,a) - Q_G(s_2,a)\right| \leq \varepsilon \left(\frac{|\mathcal{A}|}{1-\gamma} + k\varepsilon + k \right).
\label{eq:bolt_qs}
\end{equation}
Consequently, we can consider $\epB$ as a variation of $\epQ$, where the difference in $Q$-values is bounded by $\varepsilon \left(\frac{|\mathcal{A}|}{1-\gamma} + k\varepsilon + k \right)$. Lemma~\ref{lma:bolt_lemma} then follows directly from Lemma~\ref{lma:Q*}.

@@

{\bf Proof Sketch of Lemma~\ref{lma:mult_lemma}} The proof follows an identical strategy to that of Lemma~\ref{lma:bolt_lemma}, but without $e^x \approx 1+x$.

>>>>

We now define types of abstraction based on functions of state--action pairs.
\bdefn{$\ep_f$}
Given a function $f: \mathcal{S}_G \times \mathcal{A} \rightarrow \mathbb{R}$ and a fixed non-negative $\varepsilon \in \mathbb{R}$, we define $\ep_f$ as a type of approximate state-aggregation function that satisfies the following for any two ground states $s_1$ and $s_2$: 
\begin{equation}
\label{eq:phi_f}
\ep_f(s_1) = \ep_f(s_2) \rightarrow \forall_a \left|f(s_1, a) - f(s_2, a)\right| \leq \varepsilon.
\end{equation}
\edefn

@@

% Theorem 1
\begin{thm}
There exist at least four types of approximate state-aggregation functions, $\epQ$, $\epM$, $\epB$ and $\epMu$, for which the optimal policy in the abstract \ac{MDP}, applied to the ground \ac{MDP}, has suboptimality bounded polynomially in $\frac{1}{\varepsilon}$:
\begin{equation}
\forall_{s \in \mathcal{S}_G} V_G^{\pi_G^*}(s) - V_G^{\pi_{GA}}(s) \leq poly \left (\frac{1}{\varepsilon} \right).
\end{equation}
\end{thm}

@@

Note that as $t \rightarrow \infty$, $\sum_{i=0}^{t}\gamma^i \frac{2\varepsilon}{1-\gamma} \rightarrow \frac{2\varepsilon}{(1-\gamma)^2}$ by the sum of infinite geometric series and $\pi_{A,t}(s) \rightarrow \pi_{GA}$.
Thus, we conclude Lemma~\ref{lma:Q*}.
\qed
%\begin{equation*}
%\forall_{s_G \in \mathcal{S}_g} V_G(s_G) \leq  V_G^{\pi_{GA}}(s_G) + \frac{2\varepsilon}{(1-\gamma)^2}
%\end{equation*}
%\begin{equation*}
%\forall_{s_G \in \mathcal{S}_g} \left | V_G(s_G) - V_G^{\pi_{GA}}(s_G) \right | \leq  \frac{2\varepsilon}{(1-\gamma)^2}
%\end{equation*}

@@

Now, consider an approximate version of \citet{li2006towards}'s $\phi_{model}$, where states are aggregated together when their rewards and transitions are within $\varepsilon$.
\bdefn{$\epM$}
We let $\epM$ define a type of abstraction that, for fixed $\varepsilon$, satisfies:
\begin{multline}
\epM(s_1) = \epM(s_2) \rightarrow \\
\forall_a \left| \mcR_G(s_1, a) - \mcR_G(s_2, a)\right| \leq \varepsilon\; \text{~AND}\ \\
\forall_{s_A \in \mcS_A} \left|\sum_{{s_G}' \in G(s_A)} \left[\mcT_G(s_1, a, {s_G}') - \mcT_G(s_2, a,{s_G}')\right] \right| \leq \varepsilon.
\end{multline}
\edefn

@@

Let $B$ be the maximum $Q$-value difference between any pair of ground states in the same abstract state for $\epM$:
\begin{equation*}
B = \max_{s_A, s_1, s_2, a}  |Q_G(s_1, a) - Q_G(s_2, a)|,
\end{equation*}
where $s_A \in \mcS_G$ and $s_1, s_2 \in G(s_A)$. Since difference of rewards is bounded by $\varepsilon$:
% Removed for space
%\begin{multline*}
%B=\max_{s_A, s_1, s_2, a}      \biggl|\mcR_G(s_1, a) - \mcR_G(s_2, a)\ + \gamma \sum_{{s_G}' \in \mcS_G} \\
%\biggl[(\mcT_G(s_1,a,{s_G}') -\mcT_G(s_2, a, {s_G}'))\max_{a'}Q_G({s_G}', a')\biggr]\biggr|
%\end{multline*}
\begin{multline*}
B\leq \varepsilon + \gamma \sum_{s_A \in \mcS_A}\sum_{{s_G}' \in G(s_A)} \biggl[(T_G(s_1, a, {s_G}')\ -\\ 
T_G(s_2, a, {s_G}')) \max_{a'}Q_G({s_G}', a') \biggr].
\end{multline*}
By similarity of transitions under $\epM$:
\begin{align*}
B \leq \varepsilon + \gamma \textsc{QMax} \sum_{s_A \in \mcS_A} \varepsilon \leq \varepsilon + \gamma|\mcS_G|\varepsilon \textsc{QMax}.
\end{align*}
Since \textsc{QMax} $\leq \frac{\textsc{RMax}}{1-\gamma}$, and we defined $\textsc{RMax} = 1$:
\begin{equation*}
B \leq \frac{\varepsilon + \gamma(|\mcS_G| - 1) \varepsilon}{1-\gamma}.
\end{equation*}
Since the $Q$-values of ground states grouped under $\epM$ are strictly less than $B$, we can understand $\epM$ as a type of $\epQ$ with $\varepsilon = B$. Applying Lemma \ref{lma:Q*} yields Lemma \ref{lma:model}.
\qed

\subsection{Boltzmann over Optimal Q: $\epB$}
\label{sec:boltz}

@@

By the approximation in Equation~\ref{eq:e_to_x_approx} and the assumption in Equation~\ref{eq:bolt_denom}:
\begin{align*}
\left|\frac{1 + Q_G(s_1,a) + \delta_1}{\sum_j e^{Q_G(s_1,a_j)}} - \frac{1 + Q_G(s_2,a) + \delta_2}{\sum_j e^{Q_G(s_1,a_j)} \pm k\varepsilon}\right| \leq \varepsilon
\end{align*}
It follows that:
\begin{equation}
\left|Q_G(s_1,a) - Q_G(s_2,a)\right| \leq \varepsilon \left(\frac{|\mathcal{A}|}{1-\gamma} + k\varepsilon + k \right).
\label{eq:bolt_qs}
\end{equation}
Consequently, we can consider $\epB$ as a variation of $\epQ$, where the difference in $Q$-values is bounded by $\varepsilon \left(\frac{|\mathcal{A}|}{1-\gamma} + k\varepsilon + k \right)$. Lemma~\ref{lma:bolt_lemma} then follows directly from Lemma~\ref{lma:Q*}.
\qed

@@

{\bf Proof Sketch of Lemma~\ref{lma:mult_lemma}} The proof follows an identical strategy to that of Lemma~\ref{lma:bolt_lemma}, but without $e^x \approx 1+x$.
\qed
